Fishery Markets and Large-Scale Marine Conservation

### 2019-05-30 docs/old_ms/Markets_Conservation_May28.tex
Added all files

% Place your abstract within the special {sciabstract} environment.
\begin{sciabstract}
Will rights-based approaches to managing natural resources facilitate, or impede large-scale conservation? We study this question in the context of large-scale marine protected areas, and ask whether a country that commits to large-scale conservation is destined to suffer commensurate losses in fishery income.  Indeed, this is a question maritime nations all over the world face as they consider whether, and how, to comply with various international commitments to expand marine protection. We use a combination of spatial bioeconomic modeling, which accounts for fisheries markets such as the vessel-day trading scheme in the Pacific Ocean, and vessel-level empirical data on fishing behavior before and after MPA implementation, to assess how market design incentivizes, or penalizes conservation. The answer hinges critically on two design features of the rights-based fishery market. First, only if trading of harvest or vessel-days is allowed across countries can a conservation-minded country capture the economic spillover benefits of their conservation actions.  Instead, if trading is forbidden, countries are likely to suffer losses from conservation. Second, when rights are allocated between countries, the allocation must depend only on biological features, and not on economic factors such as a country's past effort or catch. Otherwise, the allocation process acts as a punishment to the conserving country. Overall, these results show that seemingly innocuous design features of a market can have indelible impacts on a country's willingness to engage in large-scale conservation; these results are confirmed with an empirical case study from the Phoenix Islands Protected Area, one of the world's largest marine reserves. With the right design features, large-scale protected areas such as the Phoenix Islands can be achieved at close to zero cost to the implementing country.
\end{sciabstract}

@@

In addition to highly productive tuna fisheries, the PNA waters provide a wealth of other ecosystem benefits, hence the focus on large-scale conservation efforts in the region. The PNA member nations derive enormous benefits from leasing tuna fishing rights to foreign fleets, in some cases exceeding 50\% of a country's annual revenue. A natural concern arises: even if the closure creates net benefits to the world, and perhaps even to tuna fisheries Pacific-wide, the economic loss to the conserving country could still be extremely large, and may prevent it from engaging in any meaningful conservation. Intuitively, then, it seems that a country's willingness to engage in large-scale conservation will depend crucially on whether it can successfully capture a sufficiently high fraction of the benefits that are derived from that conservation. We show how a rights-based fisheries market can be designed (if it is new) or modified (if it already exists) to ensure that the conserving country is rewarded, not punished, for large-scale conservation.

We develop a 10-patch spatial bioeconomic model of a fishery operating under a vessel-day scheme (VDS), and where conservation is being considered. A VDS is a type of rights-based management approach used by the PNA where each country is given an allocation of vessel-days that can then be sold to fishing firms. A vessel-day for country $i$ gives a firm the right to have one vessel fish in country $i$'s EEZ for 24 hours. Patches 1 - 9 operate under a VDS and Patch 10 represents the high seas in open access. Patch 1 considers closing a portion of its waters as a reserve ($R$, between 0 and 1). There are a total of 45,000 vessel-days, equally distributed between the 9 VDS-managed patches. The fishery revenue to each of these patches is the product of within-patch vessel-days and vessel-day price. Fishing vessels decide to fish in each patch based on available biomass (\emph{i.e.} biomass not within a reserve), and are willing to pay a vessel-day price that equates to the marginal profits of the last unit of effort applied. Total biomass is equally distributed across the 10 patches. The stock remains within each patch during the fishing season, but growth from escapement is equally distributed across the 10 patches.

We test the implications of conservation under two market designs: one in which effort trading between countries is not allowed, and one where vessel-days can be traded between countries. In both cases we solve for the equilibrium vessel-day price and effort redistribution. We quantify the losses by comparing revenues from each scenario to a benchmark scenario without any conservation action. The fishery is simulated for a range of reserve sizes and assumptions about within-patch stock movement. The second market design feature we investigate is the allocation of rights. We vary the degree to which periodic allocation is based on historical within-patch effort versus the actual abundance of fish biomass in a patch. Finally, we use vessel-tracking data and official reports of government revenue to corroborate our results by looking at the Phoenix Islands Protected Area, implemented in 2015 by Kiribati within the PNA. \textbf{We use the same data and insights from our model to hypothesize what might be the impacts of the proposed Palau National Marine Sanctuary, currently targeting an 80\% closure of its Exclusive Economic Zone \cite{cimino_2019}.}

We begin by simulating a fishery where effort is not traded between countries. We find that a spatial closure in Patch 1 will always result in a loss in revenues to Patch 1 (Fig. \ref{fig:PNA_model}A). Closing a greater proportion of the patch results in greater costs, but these decrease if within-patch stock movement ($\theta$) increases, allowing vessels to harvest the stock within the remaining open area. The closure-to-cost ratio for any reserve size is greater than 1:1 when stock movement is low (\emph{i.e.} $\theta < 0.2$), implying that a 30\% closure would result in at least a 30\% loss in revenues. Even for a highly mobile stock, a reserve reduces the amount of biomass that is available for harvest in Patch 1, which reduces a vessel's willingness to pay to fish in such waters (Fig. \ref{fig:vessel_day_price_no_trading_plot}). When countries cannot trade, the costs of conservation are large and incurred by Patch 1, but the benefits are received by the eight other patches (Figure S2 shows that the increase in revenue ranges from 0\% to 6\%).

How would a different market design, where trading between countries is allowed, change these results? To answer this we simulate the exact same fishery, but now allow for vessel-days to be traded across patches. A closure in Patch 1 lowers the value of vessel-days in that patch. Vessel-days in patches 2 - 9 are now worth more causing vessel-days from Patch 1 to be traded to Patches 2 - 9. Trading continues until vessel-day prices are the same across all patches (Fig. \ref{fig:vessel_day_price_with_trading_plot}). Under this market design, we find that revenue losses are less than 1\%, compared to the benchmark scenario with no reserve (Fig. \ref{fig:PNA_model}). This means that between 88\% and 99\% of the costs of conservation can be avoided if markets are designed to enable trading (Fig. \ref{fig:PNA_model}C).

A second concern about spatial closures in fishing effort markets is that of the allocation rules and their implications for costs and the long-term sustainability of the closures. If allocation rules updates based on the historical patch-level effort, then trading vessel-days to other countries would imply a future reduction in allocation. We simulate the fishery 50 years into the future and annually re-allocate vessel-days based on a 7-year running mean of patch-level effort and biomass. We use a range of allocation rules to weight effort ($\alpha$) and biomass ($1 - \alpha$) and compare the revenues to a fishery operating for 50 years without any closures. We find that when allocation is based on historical effort only (\emph{i.e.} $\alpha = 1$), implementing a reserve results in losses of 20\% to 93\%, depending on the size of reserve and stock movement (Fig. \ref{fig:allocation_cost_plot}). However, these costs are reduced as the weight on historical biomass is increased. A biomass-only allocation rule (\emph{i.e.} $\alpha = 0$) results in costs as low as 0.1\% to 0.7\%, essentially eliminating the costs of conservation.

We have shown that rights-based fisheries management may reduce the costs of conservation if trading is allowed, and that the costs can be further reduced if the allocation of rights is based on biomass, rather than on effort. We now move to an empirical setting where we use the implementation of one of the world's largest MPAs and vessel-tracking data to identify the effort displacement and possible costs of conservation. \textbf{We then combine our predictions and empirical results to predict the likely outcomes of upcoming large-scale MPAs in the region.}

Kiribati implemented the Phoenix Islands Protected Area in January 2015 (Fig. \ref{fig:PNA_map}). With an extension of 397,447 Km\textsuperscript{2}, this 11.5\% closure effectively displaced all fishing effort within its boundaries \cite{mccauley_2016,mcdermott_2018}. We use identification of fishing activity via Automatic Identification Systems (AIS) provided by Global Fishing Watch to track 313 tuna purse seine vessels that fished in PNA waters between 2012 and 2018 \cite{kroodsma_2018}. We continuously observe 92 vessels for the 2012 - 2018 period. Of these, 64 vessels fished within PIPA at least once prior to its implementation, and we refer to them as the ``displaced vessels''. The remaining 28 vessels never fished in PIPA waters, and we refer to them as ``non-displaced vessels''. The group with the remaining 221 vessels contains vessels that were not continuously observed before and after the implementation of PIPA, and we refer to these as ``other vessels''.

>>>>

% Place your abstract within the special {sciabstract} environment.
\begin{sciabstract}
Will rights-based approaches to managing natural resources facilitate, or impede large-scale conservation? We study this question in the context of large-scale marine protected areas, and ask whether a country that commits to large-scale conservation is destined to suffer commensurate losses in fishery income.  Indeed, this is a question maritime nations all over the world face as they consider whether, and how, to comply with international commitments to expand marine protection. We use a combination of spatial bioeconomic modeling, which accounts for fisheries markets such as the vessel-day trading scheme in the Pacific Ocean, and vessel-level empirical data on fishing behavior before and after MPA implementation, to assess how market design incentivizes, or penalizes conservation. The answer hinges critically on two design features of the rights-based fishery market. First, only if trading of harvest or vessel-days is allowed across countries can a conservation-minded country capture the economic spillover benefits of their conservation actions.  Instead, if trading is forbidden, countries are likely to suffer losses from conservation. Second, when rights are allocated between countries, the allocation must depend only on biological features, and not on economic factors such as a country's past effort or catch. Otherwise, the allocation process acts as a punishment to the conserving country. Overall, these results show that seemingly benign design features of a market can have indelible impacts on a country's willingness to engage in large-scale conservation; these results are confirmed with an empirical case study from the Phoenix Islands Protected Area, one of the world's largest marine reserves. With the right design features, large-scale protected areas such as the Phoenix Islands can be achieved at close to zero cost to the implementing country.
\end{sciabstract}

@@

In addition to highly productive tuna fisheries, the PNA waters provide a wealth of other ecosystem benefits, hence the focus on large-scale conservation efforts in the region. The PNA member nations derive enormous benefits from leasing tuna fishing rights to foreign fleets, in some cases exceeding half of a country's GDP. A natural concern arises: even if the closure creates net benefits to the world, and perhaps even to tuna fisheries Pacific-wide, the economic loss to the conserving country could still be extremely large, and may prevent it from engaging in any meaningful conservation. Intuitively, then, it seems that a country's willingness to engage in large-scale conservation will depend crucially on whether it can successfully capture a sufficiently high fraction of the benefits that are derived from that conservation. We show how a rights-based fisheries market can be designed (if it is new) or modified (if it already exists) to ensure that the conserving country is rewarded, not punished, for large-scale conservation.

To mirror the strategies and spatial connections among the nine PNA countries and the high seas, we develop a 10-patch spatial bioeconomic model. The fishery in nine of the patches operates under a market-based system in which ``vessel days'' are capped, closely tracked, and may be traded across countries.  This vessel-day scheme (VDS) is acutally used by the PNA; each country is given an allocation of vessel-days that can then be sold to fishing firms. A vessel-day owned by country $i$ entitles $i$ to lease that day to a fishing vessel, entitling it to fish in $i$'s EEZ for a 24 hour period. In our model, Patches 1 - 9 are countries that operate under a VDS (in which fishing days are limited) and Patch 10 represents the high seas, where fishing days are governed by economic conditions, not by the VDS.

While modeling the VDS market may be interesting in its own right, our primary purpose is to examine how the VDS affects a country's incentives for large-scale conservation. To examine this, suppose that Country 1 considers closing a portion $R$ (between 0 and 1) of its waters as a marine reserve, in which no fishing is allowed. When Country 1 creates a reserve, several consequences may arise.  Perhaps most obviously, the area available to fish in Country 1 is now limited, which could reduce the amount of money fishing firms are willing to pay for a day of fishing in $i$'s waters.  Second, to the extent that catch is reduced in Country 1, fish may become more abundant elsewhere; this could imply an increase in the lease price to fish in the other eight PNA countries' waters because the VDS price reflects the profit from fishing in a country, which depends crucially on the abundance of fish. For the same reason, the increase in abundance could attract new entrants on the high seas, who could quickly dissipate the benefits there. The purpose of our model is to examine these consequences in detail, with a particular focus on how design features of the market affect Country 1's incentives to create the marine reserve.

Not all market-based approaches to environmental management are created equal. A pervasive finding across a range of natural resources is that features of market design can have enormous implications for the market's functioning. In the context of fishery markets, two market design features stand out as important for a variety of outcomes, and we will show that they are pivotal in determining the incentives for large-scale marine conservation. The first design feature has to do with the trading of rights, and the second design feature concerns the allocation of rights. We examine the effects of large-scale conservation in Country 1 across a range of scenarios for trading and allocation of vessel day fishing rights across our nine model countries. In all cases we solve for the equilibrium vessel-day price, fishing effort redistribution, and fish stock a that would be expected to occur in the market. We quantify the change in revenue to all PNA countries from large-scale conservation in Country 1, by comparing revenues from each scenario to a benchmark scenario without any conservation action. We simulate these outcomes across a range of reserve sizes and assumptions about within-patch stock movement.

Why would trading and allocation rules affect the incentives for conservation by Country 1?  Invoking reduction ad absurdum, consider the incentives for Country 1 to close 100\% of its waters. Such a closure would surely benefit the other countries through the increased abundance of fish.  If Country 1 could sell its fishing rights, the other countries would likely buy them, and this could offset the foregone costs of fishing it its own waters. But if Country 1 was not allowed to sell its rights, then it would lose all of its fishing revenue.  This provides the simple intuition for why allowing trading across countries is so critical to conservation incentives. The rules for how fishing rights are allocated across the nine countries are equally important. Suppose rights are allocated each year based on the previous year's fishing - the more a country fishes, the more it gets allocated the next year.  This would clearly disadvantage a conservation-minded country.

In our bioeconomic model, we examine the implications of conservation under alternative market designs: with, and without trading, and under alternative allocation rules. To nest how the PNA allocates rights, we vary the degree to which periodic allocation is based on historical within-patch effort versus the actual abundance of fish biomass in a patch. Finally, we use vessel-tracking data and official reports of government revenue to corroborate our results by looking at the Phoenix Islands Protected Area, implemented in 2015 by Kiribati within the PNA.

We begin by simulating a fishery where trading of fishing days across countries is forbidden. In that setting, we find that a spatial closure in Patch 1 will always result in a loss in revenues to Patch 1 (Fig. \ref{fig:PNA_model}A). Closing a greater proportion of the patch results in greater costs, but these decrease if within-patch stock movement ($\theta$) increases, allowing vessels to harvest the stock within the remaining open area. The closure-to-cost ratio for any reserve size is greater than 1:1 when stock movement is low (\emph{i.e.} $\theta < 0.2$), implying that a 30\% closure would result in at least a 30\% loss in revenues. Even for a highly mobile stock, a reserve reduces the amount of biomass that is available for harvest in Patch 1, which reduces a vessel's willingness to pay to fish in such waters (Fig. \ref{fig:vessel_day_price_no_trading_plot}). When countries cannot trade, the costs of conservation are large and incurred by Patch 1, but the benefits are received by the eight other patches (Figure S2 shows that the increase in revenue ranges from 0\% to 6\%).

How would a different market design, where trading between countries is allowed, change these results? To answer this we simulate the exact same fishery, but now allow for vessel-days to be traded across patches. A closure in Patch 1 lowers the value of vessel-days in that patch. Vessel-days in patches 2 - 9 are now worth more causing vessel-days from Patch 1 to be traded to Patches 2 - 9. Trading continues until vessel-day prices are the same across all patches (Fig. \ref{fig:vessel_day_price_with_trading_plot}). Under this market design, we find that revenue losses are less than 1\%, compared to the benchmark scenario with no reserve (Fig. \ref{fig:PNA_model}). Overall, we find that between 88\% and 99\% of the costs of conservation can be avoided if markets are designed to enable trading (Fig. \ref{fig:PNA_model}C).

To analyze the consequences of different allocation rules on conservation incentives, we model an annual allocation that is a weighted function of historical fishing in a country and historical biomass of fish in a country. If allocation rules are based only on the historical effort in a country, then trading vessel-days to other countries would imply a future reduction in allocation. We simulate the fishery 50 years into the future and annually re-allocate vessel-days based on a 7-year running mean of patch-level effort and biomass. We use a range of allocation rules to weight effort ($\alpha$) and biomass ($1 - \alpha$) and compare the revenues to a fishery operating for 50 years without any closures. We find that when allocation is based on historical effort only (\emph{i.e.} $\alpha = 1$), implementing a reserve results in losses of 20\% to 93\%, depending on the size of reserve and stock movement (Fig. \ref{fig:allocation_cost_plot}). However, these costs are reduced as the weight on historical biomass is increased. A biomass-only allocation rule (\emph{i.e.} $\alpha = 0$) results in costs as low as 0.1\% to 0.7\%, essentially eliminating the costs of conservation. This result implies that if allocation is based purely on the biological features of a stock (i.e. the biomass within a country's waters), and not on fishing effort, the incentives for conservation are much greater.

We have shown that design features of a rights-based fisheries management system have enormous implications for the incentives a country faces to engage in large-scale conservation. Allowing tradeing of rights across countries, and re-allocating rights annual based on biological variables, not fishing pressure, can essentially eliminate the cost of conservation - even for very large reserves. We now move to an empirical setting where we combine vessel-tracking data and the implementation of one of the world's largest marine reserves to identify the effort displacement and likely costs of conservation.

The Pacific island nation of Kiribati implemented the Phoenix Islands Protected Area in January 2015 (Fig. \ref{fig:PNA_map}). With an extension of 397,447 Km\textsuperscript{2}, this 11.5\% closure effectively displaced all fishing effort within its boundaries \cite{mccauley_2016,mcdermott_2018}. We use identification of fishing activity via Automatic Identification Systems (AIS) provided by Global Fishing Watch to track 313 tuna purse seine vessels that fished in PNA waters between 2012 and 2018 \cite{kroodsma_2018}. We continuously observe 92 vessels for the 2012 - 2018 period. Of these, 64 vessels fished within PIPA at least once prior to its implementation, and we refer to them as the ``displaced vessels''. The remaining 28 vessels never fished in PIPA waters, and we refer to them as ``non-displaced vessels''. The group with the remaining 221 vessels contains vessels that were not continuously observed before and after the implementation of PIPA, and we refer to these as ``other vessels''.

COSTELLO STOPPED HERE
